IRS-aided SWIPT: Joint waveform, active and passive beamforming design

### 2020-09-27 main.tex
:art: Update discussion and style

\input{packages.tex}

\begin{document}

	\title{IRS-aided SWIPT: Joint waveform, active and passive beamforming design}
	\author{Yang Zhao, Bruno Clerckx}

	\maketitle

	\begin{abstract}
		The performance of Simultaneous Wireless Information and Power Transfer (SWIPT) is mainly restricted by the strength of the received Radio-Frequency (RF) signal. To tackle this problem, we introduce a low-power Intelligent Reflecting Surface (IRS) that compensates the propagation loss and boosts the energy efficiency with a passive beamforming gain. This paper investigates an IRS-aided Orthogonal Frequency Division Multiplexing (OFDM) SWIPT system based on a practical nonlinear energy harvester model, where a multi-antenna Access Point (AP) transmits information and energy simultaneously to a single-antenna user under the assist of IRS. We aim to maximize the Rate-Energy (R-E) region via jointly optimizing the transmit waveform at the AP, the reflection coefficients at the IRS, and the power splitting ratio at the user. The performance of the proposed waveform and beamforming design is compared with those of no IRS, fixed IRS and ideal frequency-selective (FS) IRS, and we confirm that due to rectifier nonlinearity, a dedicated power signal can be beneficial to energy harvesting (EH) while the optimal transceiving strategy depends on the system configuration. Simulation results also demonstrate that the proposed adaptive IRS design brings significant R-E enhancement over benchmark schemes for wideband transmission, and the optimal IRS can be approximated in closed form with negligible performance loss for narrowband transmission.
	\end{abstract}

@@

			The concept of SWIPT were first cast in \cite{Varshney2008}, where the authors investigated the Rate-Energy (R-E) tradeoff for a flat Gaussian channel and typical discrete channels. Two practical receiver structures were then proposed in \cite{Zhou2013}, namely Time Switching (TS) that switches between Energy Harvesting (EH) and Information Decoding (ID) modes, and Power Splitting (PS) that splits the received signal into individual components. On top of this, \cite{Zhang2013} characterized the R-E region for a Multiple-Input Multiple-Output (MIMO) broadcast system under TS and PS setup. Information and power beamforming was then considered in multiuser Multi-Input Single-Output (MISO) systems to maximize the Weighted Sum-Power (WSP) subjective to Signal-to-Interference-plus-Noise Ratio (SINR) constraints \cite{Xu2014}. Motivated by this, \cite{Krikidis2014} investigated fundamental transceiver modules, scheduling schemes, and interference management for SWIPT systems. However, \cite{Boshkovska2015} pointed out that the Radio Frequency-to-Direct Current (RF-to-DC) conversion efficiency depends on the harvester input power level. The authors also suggested a curve-fitting based parametric harvester model and proposed an iterative resource allocation algorithm. From another perspective, \cite{Trotter2009,Boaventura2011} demonstrated that multisine waveform is more suitable for WPT as it outperforms single tone in both operation range and RF-to-DC efficiency. \cite{Clerckx2016a} derived a tractable nonlinear harvester model based on the Taylor expansion of diode I-V characteristics, then implemented waveform design for WPT. Simulation and experiments demonstrated the benefit of modelling rectifier nonlinearity in system design \cite{Kim2019,Kim2019a}. The work was extended to SWIPT in \cite{Clerckx2018} where a superposition of modulated information waveform and multisine power waveform was employed to enlarge the R-E region. In contrast, \cite{Park2018} suggested an adaptive dual-mode SWIPT, which alternates between single-tone transmission that exploits conventional modulation for high-rate applications and multisine transmission that encodes the information in the Peak-to-Average Ratio (PAPR) for power-demanding applications. Modulation and coding schemes for SWIPT also receive much attention. By assuming On-Off-Keying (OOK) where bit 1 carries energy, \cite{Hu2019} compared unary and Run-Length-Limited (RLL) code in terms of rate versus battery overflow/underflow probability. It suggested that WPT should be activated only at constellation points with a large offset. Also, a learning approach \cite{Varasteh2019c} demonstrated that the offset of the power symbol is positively correlated to the harvester energy constraint, while the optimal information symbols are symmetrically distributed around the origin. It confirmed that the superposed waveform is effective to enlarge R-E region when considering rectifier nonlinearity. As for the network design, \cite{Liu2016} proposed a cooperative SWIPT Non-Orthogonal Multiple Access (NOMA) protocol with three user selection schemes such that the strong user assists the EH of the weak user. SWIPT based on Rate Splitting (RS) technique was also explored in \cite{Mao2019}.
		\end{subsection}

@@

		\begin{subsection}{Objective and Methodology}
			In this paper, we study an IRS-aided downlink MISO SWIPT system where the IRS assists the information and energy transmission of a single user. A multicarrier unmodulated power waveform (deterministic multisine) is superposed to a multicarrier modulated information waveform (e.g. OFDM) to boost the energy transfer efficiency without creating additional interference. The transmit waveform, IRS phase shift and receive splitting ratio are jointly optimized to maximize the R-E tradeoff. Different from previous research, this paper focus on multicarrier IRS-SWIPT and investigates the fundamental impact of harvester nonlinearity on passive beamforming design. The R-E region characterization problem is transformed into multiple current maximization problems subject to different rate constraints. To reduce the design complexity, we propose an Alternating Optimization (AO) algorithm that updates the channel and transceiver iteratively based on Semidefinite Relaxation (SDR) and Geometric Programming (GP) techniques. Numerical results showed that SDR is tight and the proposed algorithm can find a stationary point for all tested channel realizations. We demonstrate that dedicated power waveform can boost the energy transmission efficiency such that TS and PS are preferred at low SNR and high SNR, respectively. Also, IRS brings a significant channel amplification and R-E enhancement especially when deployed next to the transmitter or the receiver. Finally, the proposed adaptive IRS design outperformed the benchmark schemes for wideband transmission, and the optimal IRS can be approximated in closed form for narrowband transmission.

			\textit{Organization:} The rest of this paper is organized as follows. Section \ref{se:system_model} introduces the signal, channel, decoder, harvester, and R-E tradeoff models of the IRS-aided SWIPT system. Section \ref{se:problem_formulation} tackles the waveform, active and passive beamforming optimization. Section \ref{se:performance_evaluation} presents simulation results to evaluate the proposed design. Section [TODO] concludes the paper.

@@

		As shown in Fig. \ref{fi:system}, we consider an IRS-aided SWIPT system where a $M$-antenna AP delivers information and power simultaneously, through a $L$-reflector IRS, to a single-antenna user over $N$ orthogonal evenly-spaced subbands with center frequency $f_n$ ($n=1,\dots,N$). Perfect Channel State Information (CSI) with negligible training overhead is assumed to explore the upper bound of the proposed design. A quasi-static block fading channel model is considered for all links, and we focus on one particular block where the channels are approximately unchanged. Two practical co-located receiver architectures are compared in terms of R-E region. Specifically, TS divides each time slot into orthogonal data and energy slots and performs a time sharing between WPT and Wireless Information Transfer (WIT). In comparison, PS splits the received signal into individual ID and EH streams such that the splitting ratio $\rho$ is coupled with waveform and IRS design. Perfect synchronization is assumed among the three parties in both scenarios, and signals reflected by IRS for two and more times are omitted. We also assume the noise is too small to harvest.

@@

			In this section, we briefly revisit a tractable nonlinear rectenna model that relates the harvester output DC current to the received waveform \cite{Clerckx2016a,Clerckx2018b}. Fig. \ref{ci:antenna_equivalent_circuit} illustrates the equivalent circuit of a lossless antenna, where the incoming signal creates an voltage source $v_s(t)$ and the antenna has an impedance $R_{\text{ant}}$. Let $R_{\text{in}}$ be the total input impedance of the rectifier and matching network, and we assume the voltage across matching network is negligible. When perfectly matched ($R_{\text{in}}=R_{\text{ant}}$), the rectifier input voltage is $v_{\text{in}}(t)=y(t)\sqrt{\rho R_{\text{ant}}}$.

			Rectifiers consist of nonlinear components as diode and capacitor to produce DC output and store energy \cite{Hagerty2004,Pinuela2013}. Consider a simplified rectifier in Fig. \ref{ci:single_diode_rectifier} where a single series diode is followed by a low-pass filter with a parallel load. Denote $i_s$ as the reverse bias saturation current, $n'$ as the diode ideality factor, $v_t$ as the thermal voltage, $v_d(t)=v_{\text{in}}(t)-v_{\text{out}}(t)$ as the voltage across the diode where $v_{\text{out}}(t)$ is the output voltage across the load. A Taylor expansion of the diode characteristic equation $i_d(t)=i_s(e^{v_d(t)/n' v_t}-1)$ around a quiescent operating point $a$ writes as $i_d(t)=\sum_{i=0}^{\infty}k_i'(v_d(t)-a)^i$, where $k_0'=i_s(e^{a/n' v_t}-1)$ and $k_i'=i_se^{a/n'v_t}/i!(n'v_t)^i$ for $i=1,\dots,\infty$. Note that this small-signal expansion model is only valid for the non-linear operation region, and the I-V relationship would be linear if the diode behavior is dominated by the load \cite{Clerckx2016a}. Also, an ideal low-pass filter with steady-state response can provide a constant $v_{\text{out}}$ that depends on the peak of $v_{\text{in}}(t)$ \cite{Curty2005}. Therefore, a proper choice of the operating voltage drop is $a=\mathcal{E}\{v_d(t)\}=-v_{\text{out}}$ such that
			\begin{equation}\label{eq:i_d}
				i_d(t)=\sum_{i=0}^{\infty}k_i'\rho^{i/2}R_{\text{ant}}^{i/2}y(t)^i
			\end{equation}

@@

		\begin{subsection}{Rate-Energy Region}
			Define the achievable R-E region as
			\begin{align}
				C_{R_{\text{ID}}-I_{\text{EH}}}(P)
				&\triangleq \biggl\{(R_{\text{ID}}, I_{\text{EH}}): R_{\text{ID}} \le R, I_{\text{EH}} \le z,\nonumber\\
				&\quad \frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right) \le P\biggr\}
			\end{align}
			where $P$ is the average transmit power budget and the coefficient $1/2$ converts the peak power of sine waves to the average power.
		\end{subsection}
	\end{section}

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through multiple current maximization problems subject to transmit power, IRS magnitude, and different rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1}
		\end{maxi!}
		Problem \ref{op:original} is intricate due to coupled variables involved in non-convex objective function \ref{ob:original} and rate constraint \ref{co:original_rate}. To reduce the design complexity, we propose an suboptimal AO algorithm that iteratively updates the IRS phase shift and transmit waveform plus the receive splitting ratio until convergence.

		\begin{subsection}{IRS Phase Shift}
			In this section, the IRS phase shift $\boldsymbol{\phi}$ is optimized for any given waveform $\boldsymbol{w}_{I/P}$ and splitting ratio $\rho$. We observe that
			\begin{align}
				\lvert \boldsymbol{h}_{n}^H\boldsymbol{w}_{I,n} \rvert^2
				& = \boldsymbol{w}_{I,n}^H\boldsymbol{h}_n\boldsymbol{h}_n^H\boldsymbol{w}_{I,n}\nonumber\\
				& = \boldsymbol{w}_{I,n}^H(\boldsymbol{h}_{D,n}+\boldsymbol{V}_n^H\boldsymbol{\phi})(\boldsymbol{h}_{D,n}^H+\boldsymbol{\phi}^H\boldsymbol{V}_n)\boldsymbol{w}_{I,n}\nonumber\\
				& = \boldsymbol{w}_{I,n}^H\boldsymbol{M}_n^H\boldsymbol{\Phi}\boldsymbol{M}_n\boldsymbol{w}_{I,n}\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{M}_n\boldsymbol{w}_{I,n}\boldsymbol{w}_{I,n}^H\boldsymbol{M}_n^H\boldsymbol{\Phi})\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{C}_n\boldsymbol{\Phi})
			\end{align}
			where $t$ is an auxiliary variable with unit modulus, $\boldsymbol{M}_n=[\boldsymbol{V}_n^H, \boldsymbol{h}_{D,n}]^H \in \mathbb{C}^{(L+1) \times M}$, $\bar{\boldsymbol{\phi}}=[\boldsymbol{\phi}^H, t]^H \in \mathbb{C}^{(L+1) \times 1}$, $\boldsymbol{\Phi}=\bar{\boldsymbol{\phi}}\bar{\boldsymbol{\phi}}^H \in \mathbb{C}^{(L+1) \times (L+1)}$, $\boldsymbol{C}_n = \boldsymbol{M}_n\boldsymbol{w}_{I,n}\boldsymbol{w}_{I,n}^H\boldsymbol{M}_n^H \in \mathbb{C}^{(L+1)\times(L+1)}$. Also, define $t_{I/P,n}$ ($n=-N+1,\dots,N-1$) as
			\begin{align}
				t_{I/P,n}
				& = \boldsymbol{h}^H\boldsymbol{W}_{I/P,n}\boldsymbol{h}\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{h}\boldsymbol{h}^H\boldsymbol{W}_{I/P,n})\nonumber\\
				& = \mathrm{Tr}\left((\boldsymbol{h}_{D}+\boldsymbol{V}^H\boldsymbol{\phi})(\boldsymbol{h}_{D}^H+\boldsymbol{\phi}^H\boldsymbol{V})\boldsymbol{W}_{I/P,n}\right)\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{M}^H\boldsymbol{\Phi}\boldsymbol{M}\boldsymbol{W}_{I/P,n})\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{M}\boldsymbol{W}_{I/P,n}\boldsymbol{M}^H\boldsymbol{\Phi})\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{C}_{I/P,n}\boldsymbol{\Phi})
			\end{align}
			where $\boldsymbol{V}=[\boldsymbol{V}_1,\dots,\boldsymbol{V}_N] \in \mathbb{C}^{L \times MN}$, $\boldsymbol{M}=[\boldsymbol{V}^H, \boldsymbol{h}_{D}]^H \in \mathbb{C}^{(L+1) \times MN}$, $\boldsymbol{C}_{I/P,n}=\boldsymbol{M}\boldsymbol{W}_{I/P,n}\boldsymbol{M}^H \in \mathbb{C}^{(L+1)\times(L+1)}$. Therefore, the rate and objective expressions rewrite as
			\begin{align}
				R(\boldsymbol{\Phi})
				& = \sum_{n}{\log_2\left(1+\frac{(1-\rho)\mathrm{Tr}(\boldsymbol{C}_n\boldsymbol{\Phi})}{\sigma_n^2}\right)}\label{eq:R_irs}\\
				z(\boldsymbol{\Phi})
				& = \frac{1}{2}{\beta_2}{\rho}(t_{I,0}+t_{P,0})\nonumber\\
				& \quad + \frac{3}{8}{\beta_4}{\rho^2} \left(2t_{I,0}^2 + \sum_{n=-N+1}^{N-1}{t_{P,n}t_{P,n}^*}\right)\nonumber\\'
				& \quad + \frac{3}{2}{\beta_4}{\rho^2}t_{I,0}t_{P,0}\label{eq:z_irs}
			\end{align}

>>>>

\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{array}
% \usepackage{caption}
\usepackage{cite}
\usepackage{environ}
\usepackage{grffile}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage{pgfplots}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage{url}
\usepackage{xcolor}
\usepackage[RPvoltages]{circuitikz}
\usepackage[T1]{fontenc}
\usepackage[caption=false,font=footnotesize]{subfig}
% \usepackage[cmintegrals]{newtxmath}
\usepackage[short]{optidef}

\interdisplaylinepenalty=2500
\pgfplotsset{compat=newest}
\usetikzlibrary{plotmarks}
\usetikzlibrary{arrows.meta}
\usepgfplotslibrary{patchplots}
\newtheorem{proposition}{Proposition}
\DeclareSIUnit{\belm}{Bm}
\DeclareSIUnit{\dBm}{\deci\belm}
\DeclareSIUnit{\beli}{Bi}
\DeclareSIUnit{\dBi}{\deci\beli}

\makeatletter
\newcommand{\forcealgorithm}{\let\@latex@error\@gobble}
\makeatother

\begin{document}
	\title{IRS-aided SWIPT: Joint waveform, active and passive beamforming design}
	\author{
		\IEEEauthorblockN{
			Yang~Zhao,~\IEEEmembership{Student~Member,~IEEE,}
			Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE,}
			and~Zhenyuan~Feng,~\IEEEmembership{Student~Member,~IEEE}
		}
		\thanks{
			The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: yang.zhao18@imperial.ac.uk; b.clerckx@imperial.ac.uk).
		}
	}
	\maketitle

	\begin{abstract}
		The performance of Simultaneous Wireless Information and Power Transfer (SWIPT) is mainly restricted by the strength of the received Radio-Frequency (RF) signal. To tackle this problem, we introduce a low-power Intelligent Reflecting Surface (IRS) that compensates the propagation loss and boosts the energy efficiency with a passive beamforming gain. This paper investigates an IRS-aided Orthogonal Frequency Division Multiplexing (OFDM) SWIPT system based on a practical nonlinear energy harvester model, where a multi-antenna Access Point (AP) transmits information and energy simultaneously to a single-antenna user under the assist of IRS. We aim to maximize the Rate-Energy (R-E) region via jointly optimizing the transmit waveform at the AP, the reflection coefficients at the IRS, and the power splitting ratio at the user. The performance of the proposed waveform and beamforming design is compared with those of no IRS, fixed IRS and ideal frequency-selective (FS) IRS, and we confirm that due to rectifier nonlinearity, a dedicated power signal can be beneficial to energy harvesting (EH) while the optimal transceiving strategy depends on the system configuration. Simulation results also demonstrate that the proposed adaptive IRS design brings significant R-E enhancement over benchmark schemes for broadband transmission, and the optimal IRS can be approximated in closed form with negligible performance loss for SISO narrowband transmission.
	\end{abstract}

@@

			The concept of SWIPT were first cast in \cite{Varshney2008}, where the authors investigated the Rate-Energy (R-E) tradeoff for a flat Gaussian channel and typical discrete channels. Two practical receiver structures were then proposed in \cite{Zhou2013}, namely Time Switching (TS) that switches between Energy Harvesting (EH) and Information Decoding (ID) modes, and Power Splitting (PS) that splits the received signal into individual components. On top of this, \cite{Zhang2013} characterized the R-E region for a Multiple-Input Multiple-Output (MIMO) broadcast system under TS and PS setup. Information and power beamforming was then considered in multiuser Multi-Input Single-Output (MISO) systems to maximize the Weighted Sum-Power (WSP) subjective to Signal-to-Interference-plus-Noise Ratio (SINR) constraints \cite{Xu2014}. Motivated by this, \cite{Krikidis2014} investigated fundamental transceiver modules, scheduling schemes, and interference management for SWIPT systems. However, \cite{Boshkovska2015} pointed out that the Radio Frequency-to-Direct Current (RF-to-DC) conversion efficiency depends on the harvester input power level. The authors also suggested a curve-fitting based parametric harvester model and proposed an iterative resource allocation algorithm. From another perspective, \cite{Trotter2009,Boaventura2011} demonstrated that multisine waveform is more suitable for WPT as it outperforms single tone in both operation range and RF-to-DC efficiency. \cite{Clerckx2016a} derived a tractable nonlinear harvester model based on the Taylor expansion of diode I-V characteristics, then implemented waveform design for WPT. Simulation and experiments demonstrated the benefit of modelling rectifier nonlinearity in system design \cite{Kim2019,Kim2019a}. The work was extended to SWIPT in \cite{Clerckx2018} where a superposition of modulated information waveform and multisine power waveform was employed to enlarge the R-E region. In contrast, \cite{Park2018} suggested an adaptive dual-mode SWIPT, which alternates between single-tone transmission that exploits conventional modulation for high-rate applications and multisine transmission that encodes the information in the Peak-to-Average Ratio (PAPR) for power-demanding applications. Modulation and coding schemes for SWIPT also receive much attention. By assuming On-Off-Keying (OOK) where bit \num{1} carries energy, \cite{Hu2019} compared unary and Run-Length-Limited (RLL) code in terms of rate versus battery overflow/underflow probability. It suggested that WPT should be activated only at constellation points with a large offset. Also, a learning approach \cite{Varasteh2019c} demonstrated that the offset of the power symbol is positively correlated to the harvester energy constraint, while the optimal information symbols are symmetrically distributed around the origin. It confirmed that the superposed waveform is effective to enlarge R-E region when considering rectifier nonlinearity. As for the network design, \cite{Liu2016} proposed a cooperative SWIPT Non-Orthogonal Multiple Access (NOMA) protocol with three user selection schemes such that the strong user assists the EH of the weak user. SWIPT based on Rate Splitting (RS) technique was also explored in \cite{Mao2019}.
		\end{subsection}

@@

		\begin{subsection}{Objective and Methodology}
			In this paper, we study an IRS-aided downlink MISO SWIPT system where the IRS assists the information and energy transmission of a single user. A multicarrier unmodulated power waveform (deterministic multisine) is superposed to a multicarrier modulated information waveform (e.g. OFDM) to boost the energy transfer efficiency without creating additional interference. The transmit waveform, IRS phase shift and receive splitting ratio are jointly optimized to maximize the R-E tradeoff. Different from previous research, this paper focus on multicarrier IRS-SWIPT and investigates the fundamental impact of harvester nonlinearity on passive beamforming design. The R-E region characterization problem is transformed into multiple current maximization problems subject to different rate constraints. To reduce the design complexity, we propose an Alternating Optimization (AO) algorithm that updates the channel and transceiver iteratively based on Semidefinite Relaxation (SDR) and Geometric Programming (GP) techniques. Numerical results showed that SDR is tight and the proposed algorithm can find a stationary point for all tested channel realizations. We demonstrate that dedicated power waveform can boost the energy transmission efficiency such that TS and PS are preferred at low SNR and high SNR, respectively. Also, IRS brings a significant channel amplification and R-E enhancement especially when deployed next to the transmitter or the receiver. Finally, the proposed adaptive IRS design outperformed the benchmark schemes for broadband transmission, and the optimal IRS can be approximated in closed form for narrowband SISO transmission.

			\textit{Organization:} The rest of this paper is organized as follows. Section~\ref{se:system_model} introduces the signal, channel, decoder, harvester, and R-E tradeoff models of the IRS-aided SWIPT system. Section~\ref{se:problem_formulation} tackles the waveform, active and passive beamforming optimization. Section~\ref{se:performance_evaluation} presents simulation results to evaluate the proposed design. Section~[TODO] concludes the paper.

@@

		As shown in Fig.~\ref{fi:system}, we consider an IRS-aided SWIPT system where a $M$-antenna AP delivers information and power simultaneously, through a $L$-reflector IRS, to a single-antenna user over $N$ orthogonal evenly-spaced subbands with center frequency $f_n$ ($n=1,\dots,N$). Perfect Channel State Information (CSI) with negligible training overhead is assumed to explore the upper bound of the proposed design. A quasi-static block fading channel model is considered for all links, and we focus on one particular block where the channels are approximately unchanged. Two practical co-located receiver architectures are compared in terms of R-E region. Specifically, TS divides each time slot into orthogonal data and energy slots and performs a time sharing between WPT and Wireless Information Transfer (WIT). In comparison, PS splits the received signal into individual ID and EH streams such that the splitting ratio $\rho$ is coupled with waveform and IRS design. Perfect synchronization is assumed among the three parties in both scenarios, and signals reflected by IRS for two and more times are omitted. We also assume the noise is too small to harvest.

@@

			In this section, we briefly revisit a tractable nonlinear rectenna model that relates the harvester output DC current to the received waveform \cite{Clerckx2016a,Clerckx2018b}. Fig.~\ref{ci:antenna_equivalent_circuit} illustrates the equivalent circuit of a lossless antenna, where the incoming signal creates an voltage source $v_s(t)$ and the antenna has an impedance $R_{\text{ant}}$. Let $R_{\text{in}}$ be the total input impedance of the rectifier and matching network, and we assume the voltage across matching network is negligible. When perfectly matched ($R_{\text{in}}=R_{\text{ant}}$), the rectifier input voltage is $v_{\text{in}}(t)=y(t)\sqrt{\rho R_{\text{ant}}}$.

			Rectifiers consist of nonlinear components as diode and capacitor to produce DC output and store energy \cite{Hagerty2004,Pinuela2013}. Consider a simplified rectifier in Fig.~\ref{ci:single_diode_rectifier} where a single series diode is followed by a low-pass filter with a parallel load. Denote $i_s$ as the reverse bias saturation current, $n'$ as the diode ideality factor, $v_t$ as the thermal voltage, $v_d(t)=v_{\text{in}}(t)-v_{\text{out}}(t)$ as the voltage across the diode where $v_{\text{out}}(t)$ is the output voltage across the load. A Taylor expansion of the diode characteristic equation $i_d(t)=i_s(e^{v_d(t)/n' v_t}-1)$ around a quiescent operating point $a$ writes as $i_d(t)=\sum_{i=0}^{\infty}k_i'(v_d(t)-a)^i$, where $k_0'=i_s(e^{a/n' v_t}-1)$ and $k_i'=i_se^{a/n'v_t}/i!(n'v_t)^i$ for $i=1,\dots,\infty$. Note that this small-signal expansion model is only valid for the non-linear operation region, and the I-V relationship would be linear if the diode behavior is dominated by the load \cite{Clerckx2016a}. Also, an ideal low-pass filter with steady-state response can provide a constant $v_{\text{out}}$ that depends on the peak of $v_{\text{in}}(t)$ \cite{Curty2005}. Therefore, a proper choice of the operating voltage drop is $a=\mathcal{E}\{v_d(t)\}=-v_{\text{out}}$ such that
			\begin{equation}\label{eq:i_d}
				i_d(t)=\sum_{i=0}^{\infty}k_i'\rho^{i/2}R_{\text{ant}}^{i/2}y(t)^i
			\end{equation}

@@

		\begin{subsection}{Rate-Energy Region}
			Define the achievable R-E region as
			\begin{align}
				C_{R_{\text{ID}}-I_{\text{EH}}}(P)
				&\triangleq \biggl\{(R_{\text{ID}}, I_{\text{EH}}): R_{\text{ID}} \le R, I_{\text{EH}} \le z,\nonumber\\
				&\quad \frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right) \le P\biggr\}
			\end{align}
			where $P$ is the average transmit power budget and the coefficient \num{1/2} converts the peak power of sine waves to the average power.
		\end{subsection}
	\end{section}

	\begin{section}{Problem Formulation}\label{se:problem_formulation}
		We characterize the R-E region through multiple current maximization problems subject to transmit power, IRS magnitude, and different rate constraints
		\begin{maxi!}
			{\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho}{z(\boldsymbol{\phi},\boldsymbol{w}_I,\boldsymbol{w}_P,\rho)}{\label{op:original}}{\label{ob:original}}
			\addConstraint{\frac{1}{2}\left(\lVert{\boldsymbol{w}_I}\rVert^2+\lVert{\boldsymbol{w}_P}\rVert^2\right)\le{P}}\label{co:original_power}
			\addConstraint{R(\boldsymbol{\phi},\boldsymbol{w}_I,\rho) \ge \bar{R}}\label{co:original_rate}
			\addConstraint{\lvert{\phi_l}\rvert=1, \quad l=1,\dots,L}\label{co:original_modulus}
			\addConstraint{0 \le \rho \le 1}
		\end{maxi!}
		Problem~\ref{op:original} is intricate due to coupled variables involved in non-convex objective function \ref{ob:original} and rate constraint \ref{co:original_rate}. To reduce the design complexity, we propose an suboptimal AO algorithm that iteratively updates the IRS phase shift and transmit waveform plus the receive splitting ratio until convergence.

		\begin{subsection}{IRS Phase Shift}
			In this section, the IRS phase shift $\boldsymbol{\phi}$ is optimized for any given waveform $\boldsymbol{w}_{I/P}$ and splitting ratio $\rho$. We observe that
			\begin{align}
				\lvert \boldsymbol{h}_{n}^H\boldsymbol{w}_{I,n} \rvert^2
				& = \boldsymbol{w}_{I,n}^H\boldsymbol{h}_n\boldsymbol{h}_n^H\boldsymbol{w}_{I,n}\nonumber\\
				& = \boldsymbol{w}_{I,n}^H(\boldsymbol{h}_{D,n}+\boldsymbol{V}_n^H\boldsymbol{\phi})(\boldsymbol{h}_{D,n}^H+\boldsymbol{\phi}^H\boldsymbol{V}_n)\boldsymbol{w}_{I,n}\nonumber\\
				& = \boldsymbol{w}_{I,n}^H\boldsymbol{M}_n^H\boldsymbol{\Phi}\boldsymbol{M}_n\boldsymbol{w}_{I,n}\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{M}_n\boldsymbol{w}_{I,n}\boldsymbol{w}_{I,n}^H\boldsymbol{M}_n^H\boldsymbol{\Phi})\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{C}_n\boldsymbol{\Phi})
			\end{align}
			where $t$ is an auxiliary variable with unit modulus, $\boldsymbol{M}_n=[\boldsymbol{V}_n^H, \boldsymbol{h}_{D,n}]^H \in \mathbb{C}^{(L+1) \times M}$, $\bar{\boldsymbol{\phi}}=[\boldsymbol{\phi}^H, t]^H \in \mathbb{C}^{(L+1) \times 1}$, $\boldsymbol{\Phi}=\bar{\boldsymbol{\phi}}\bar{\boldsymbol{\phi}}^H \in \mathbb{C}^{(L+1) \times (L+1)}$, $\boldsymbol{C}_n = \boldsymbol{M}_n\boldsymbol{w}_{I,n}\boldsymbol{w}_{I,n}^H\boldsymbol{M}_n^H \in \mathbb{C}^{(L+1)\times(L+1)}$. Also, define $t_{I/P,n}$ ($n=-N+1,\dots,N-1$) as
			\begin{align}
				t_{I/P,n}
				& = \boldsymbol{h}^H\boldsymbol{W}_{I/P,n}\boldsymbol{h}\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{h}\boldsymbol{h}^H\boldsymbol{W}_{I/P,n})\nonumber\\
				& = \mathrm{Tr}\left((\boldsymbol{h}_{D}+\boldsymbol{V}^H\boldsymbol{\phi})(\boldsymbol{h}_{D}^H+\boldsymbol{\phi}^H\boldsymbol{V})\boldsymbol{W}_{I/P,n}\right)\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{M}^H\boldsymbol{\Phi}\boldsymbol{M}\boldsymbol{W}_{I/P,n})\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{M}\boldsymbol{W}_{I/P,n}\boldsymbol{M}^H\boldsymbol{\Phi})\nonumber\\
				& = \mathrm{Tr}(\boldsymbol{C}_{I/P,n}\boldsymbol{\Phi})
			\end{align}
			where $\boldsymbol{V}=[\boldsymbol{V}_1,\dots,\boldsymbol{V}_N] \in \mathbb{C}^{L \times MN}$, $\boldsymbol{M}=[\boldsymbol{V}^H, \boldsymbol{h}_{D}]^H \in \mathbb{C}^{(L+1) \times MN}$, $\boldsymbol{C}_{I/P,n}=\boldsymbol{M}\boldsymbol{W}_{I/P,n}\boldsymbol{M}^H \in \mathbb{C}^{(L+1)\times(L+1)}$. Therefore, the rate and objective expressions rewrite as
			\begin{align}
				R(\boldsymbol{\Phi})
				& = \sum_{n}{\log_2\left(1+\frac{(1-\rho)\mathrm{Tr}(\boldsymbol{C}_n\boldsymbol{\Phi})}{\sigma_n^2}\right)}\label{eq:R_irs}\\
				z(\boldsymbol{\Phi})
				& = \frac{1}{2}{\beta_2}{\rho}(t_{I,0}+t_{P,0})\nonumber\\
				& \quad + \frac{3}{8}{\beta_4}{\rho^2} \left(2t_{I,0}^2 + \sum_{n=-N+1}^{N-1}{t_{P,n}t_{P,n}^*}\right)\nonumber\\
				& \quad + \frac{3}{2}{\beta_4}{\rho^2}t_{I,0}t_{P,0}\label{eq:z_irs}
			\end{align}
